# STAT 440 Statistical Data Management - Spring 2022
## Week15 Lab03 Section 01 Mon
### Both .Rmd and .html files must be submitted by the end of class plus a 10 minute grace period. Any lab submitted after that 10 minute grace period will be deducted 2 points for being late. Labs submitted after the lab solution is posted will not be accepted; resulting in a 0 for that driver's lab.
#### Only the driver should submit this assignment into their individual student repo. Save the filename as lab03-week15-section01-mon-netID.Rmd and render this file to .html. 


**#1** This problem includes concepts from the following notes: Markdown. 

**Problem:** Using Markdown syntax, make a numbered list with the netIDs of everyone at your station in lowercase text. Your netID should appear in bold.

**Answer:** Answers will vary. A numbered list includes numbers followed by period. Sample below.

1. drivernetid

2. passengernetid

**#2** This problem includes concepts from the following notes: Accessing and importing data.

**Problem:** Import the Rental Inspection Grades Listing Data using its data URL https://uofi.box.com/shared/static/lbqe1l9a2g5lr0twk6zcwajasybekget.json and print the resulting first 10 rows. The data description (included below) may be helpful.  

Data description for Rental Inspection Grades Listing Data: The dataset (a .json file) contains the 1738 rental properties in Urbana with 7 variables ("columns") per property which includes details about the parcel number, address, grade, inspection date, and more. The rental properties are inspected on a rotating basis and can receive a letter grade of A, B, C, D, F, and N. The best grade is A. The original source is the City of Urbana.

**Answer:**

```{r}
library(tidyverse)
library(jsonlite)
library(sqldf)
rentals <- fromJSON("https://uofi.box.com/shared/static/lbqe1l9a2g5lr0twk6zcwajasybekget.json")
slice_head(rentals, n=10)
```

**#3** This problem includes concepts from the following notes/videos: Arranging data, Data reduction, Data expansion, Vectorization, Regular expressions and string manipulation, Summarizing data, Combining data, SQL for weeks 03-07 content, SQL sub-queries, Tool-making and programming Part01.

**Problem:** Combine the following data about the owners of the rental properties using its data URL https://uofi.box.com/shared/static/7m2hlut865herv9zzgzvyl4ort0j7rhr.csv such that the combined data has the original 7 columns from the Rentals Data and 4 new columns: `Owner Name`, `Owner Address`, `Owner City`, and `Owner State`. Now print the last 5 columns of the new data and the following rows: 57, 120, 628, 713, 1462. **You do not have to correct any incorrectly assigned owner names, addresses, cities, or states.**

**Answer:**

```{r}
owners_addresses <- read_csv("https://uofi.box.com/shared/static/7m2hlut865herv9zzgzvyl4ort0j7rhr.csv")

own2<-str_split(owners_addresses$name, "\\\n", simplify=TRUE)
own3<-str_split(own2[,3], "\\,\\s", simplify = TRUE)
own4<-str_split(own2[,4], "\\,\\s", simplify=TRUE)

rentals2 <- bind_cols(rentals,owners_addresses) %>%
  mutate(`Owner Name` = own2[,1], `Owner Address` = own2[,2], `Owner City` = own3[,1], `Owner State` = own3[,2]) %>%
  select(-name)

rentals2 %>% 
  select(`Expiration Date`:`Owner State`) %>%
  slice(57, 120, 628, 713, 1462)
```

**#4** This problem includes concepts from the following notes/videos: Arranging data, Data reduction, Data expansion, Vectorization, Regular expressions and string manipulation, Summarizing data, Combining data, Data validating and cleaning, SQL for weeks 03-07 content, SQL sub-queries, Tool-making and programming Part01.

**Problem:** The combined data in the previous problem contains several incorrectly assigned owner names, addresses, cities, and states. **There are roughly 190 rows with incorrect values.** Validate and clean those incorrect values. Now print the last 5 columns of the newly cleaned data and the following rows: 57, 120, 628, 713, 1462.

**Answer:**

```{r}
x=ifelse(!str_detect(rentals2$`Owner Address`, "\\d"), paste(rentals2$`Owner Name`,"&",rentals2$`Owner Address`), rentals2$`Owner Name`)

xx=ifelse(!str_detect(rentals2$`Owner Address`, "\\d"), rentals2$`Owner City`, rentals2$`Owner Address`)
#sum(str_detect(xx, "\\d"))
xx2=ifelse(str_detect(xx, "^SERIES"), own3[,1], xx)

i1=which(!str_detect(xx, "^\\d"))
i2=which(!str_detect(xx[i1], "^PO"))
#xx[i1[i2]]
j1=which(!str_detect(xx2, "^\\d"))
j2=which(!str_detect(xx2[j1], "^PO"))
#j2
#xx2[j1[j2]]

#which(!str_detect(rentals2$`Owner State`, "\\w\\w")|str_detect(rentals2$`Owner State`, "\\d"))
xxx=ifelse(!str_detect(rentals2$`Owner State`, "\\w\\w")|str_detect(rentals2$`Owner State`, "\\d"), own4[,1], rentals2$`Owner City`)
#table(xxx)

xxxx=ifelse(!str_detect(rentals2$`Owner State`, "\\w\\w")|str_detect(rentals2$`Owner State`, "\\d"), own4[,2], rentals2$`Owner State`)
#table(xxxx)
#sum(str_detect(xxxx, "\\w\\w"))

rentals3 <- rentals2 %>%
  mutate(`Owner Name`= x, `Owner Address`= xx2, `Owner City`= xxx,`Owner State` = xxxx)

rentals3 %>% 
  select(`Expiration Date`:`Owner State`) %>%
  slice(57, 120, 628, 713, 1462)
```

**#5** This problem includes concepts from the following notes/videos: Data workers and their responsibilities.

**Problem:** Would you describe the skill and knowledge required to complete the above problems as those of a data engineer, data architect, or data dataist? Write your answer below in the **Answer**section following Markdown syntax.

**Answer:** Data engineer
